The development of children's ability to track and predict turn structure in conversation

### 2015-12-04 paper/CasillasFrank-Anticip.tex
updated text for submission; some small changes to go

\usepackage{graphicx}
\usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{subfig}
\usepackage{tipa}
\usepackage{float}
\usepackage{longtable}

@@

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking: early non-verbal interactions likely establish the protocol by which children come to use language with others. How do children integrate linguistic knowledge with these preverbal turn-taking abilities?

% And as children acquire language, they also come to integrate it into their preverbal turn-taking systems.

@@

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').

@@

% \subsection{}

@@

The majority of work on children's early turn taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: they take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern. A recent study by Hilbrink et al. (2015) finds that infants' turn timing at three months is often too early or too late: They start vocalizing in overlap on $~$40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average inter-turn silent gap of 350--900 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening in overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of more than 1000 msec at nine months, with only very gradual improvement after that \citep{hilbrink2015}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}.

This puzzling pattern is likely due to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance. Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}. Consistent with this hypotheiss, during the first year, children's prosodic abilities are relatively sophisticated while their lexical knowledge is limited. Infatns can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they 
% Beginning between four and five months, infants prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units \citep{jusczyk1995, soderstrom2003}. 
show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs. trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}. In comparison, children show at best a very limited lexical inventory before the first birthday \citep{bergelson2013, shi2010}. 
% Two-month-olds also notice changes in word order, but this ability appears to rely on prosodic cueing \citep{mandel1996}. 
% Generally speaking then, our current knowledge about children's linguistic development points to a possible early advantage for prosody in children's turn-taking predictions.

If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves. This hypothesis has been recently explored in experimental settings, but results are mixed. One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}. However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared. A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure. Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.

@@

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in a study of 6-, 12-, 24-, and 36-month-olds. They asked participants to watch short videos of conversation and tracked their eye movements at points of speaker change. They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was also the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: One normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 36 months), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work, it has its own limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random. Instead, they used gaze switches during ongoing speech as a baseline. Ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}, thus maximizing chances of finding a difference between gaze frequency at turn transitions and their baseline rate. A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions. We follow this conservative approach in our work.

@@

Our goal in the current study is to understand the extent to which FILL ME IN BY DESCRIBING WHAT IS UNKNOWN BEFORE OUR STUDY, AND SUMMARIZE WHAT OUR EXPERIMENTS DO FOR THAT GOAL.

% We report here on the role of linguistic processing in children's predictions about upcoming turn structure, in particular on how children use prosodic and lexicosyntactic information to make their predictions.

@@

After recording, we combined the audio and video files by hand, and cropped each recording to the 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language condition (English vs. non-English), inter-turn gap duration (in milliseconds), and speech act (question vs. non-question). The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: we asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question. We then coded turns with at least 80\% ``yes'' responses as questions.

@@

Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively). They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: They gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}). Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not. Children looked at the current speaker less than adults did during the non-English videos. Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure. Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language.

@@

@@

We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}. We built one model each for children and adults. We modeled children and adults separately because effects of age are only pertinent to the children's data. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors. Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}. We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age (3, 4, 5), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes. We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included condition, transition type, duration, and their interactions as predictors with participant and item included as random effects and random slopes of condition, transition type, and their interaction for participant.

@@

We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows. This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}). Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition analysis windows. The windows were $\pm$300 msec around the inter-turn gap.} 
\label{fig:shuffling}
\end{figure}

@@

We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data. For those interested, plots of the beta and standard error distributions are available in the Appendix.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test. For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p} $=$ .001). This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.

\begin{figure}[h!]
  \centering
  \subfloat[Children]{\includegraphics[width=0.49\textwidth]{figures/E1-chi-randrun-t-vals-absolute.png}\label{fig:E1-ChiTs}}
  \hfill
  \subfloat[Adults]{\includegraphics[width=0.49\textwidth]{figures/E1-adu-randrun-t-vals-absolute.png}\label{fig:E1-AduTs}}
  \caption{Random-permutation and original $|$\textit{t}-values$|$ for predictors of children and adults' anticipatory gaze rates. Gray dots $=$ random model estimates, White dots $=$ original model estimates, Triangles $=$ 95th percentile for each \textit{t}-value distribution.}
\end{figure}

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. The children's data showed strong evidence of differentiation from the randomly permuted data for all four significant effects in the original model (Table \ref{tab:E1-models}: Children):  the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively; Figure \ref{fig:E1-ChiTs}). Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values; Figure \ref{fig:E1-AduTs}). The effects of language condition and transition type for the real and randomly permuted data can also be observed in Figure \ref{fig:E1-randvsreal}. We excluded the output of random-permutation models that did not ultimately converge to remove unreliable model results from our percentile calculations below (78\% and 76\% of models for children and adults, respectively).

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
\end{center}
\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted (gray) data. Vertical bars represent the standard error.} 
\label{fig:E1-randvsreal}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-child-randvsreal-ttest-agebylg.png}
\end{center}
\caption{Pairwise comparisons of the language condition effect across ages for the original data (black dots) and the 5,000 randomly permuted datasets (distribution).} 
\label{fig:E1-lgageinteraction}
\end{figure}

\paragraph{Developmental effects}
The model of the children's data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have derived from random looking (Figure \ref{fig:E1-randvsreal}). To further explore this effect, we compared the average effect of language condition across all ages: we extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data. For each random permutation, we then made pairwise comparisons of the average difference scores across ages. Figure \ref{fig:E1-lgageinteraction} plots the real-data difference scores against the random-data difference score distribution for each pairwise age comparison, showing that 3- and 4-year olds were affected equally by language condition, but that 5-year-olds affected less than both 3- and 4-year-olds (with 99.52\% and 99.96\% of difference scores greater than the randomly permuted data, respectively, i.e., differences of \textit{p} $<$ .01 and \textit{p} $<$ .001).

@@

Language condition (English vs. non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}). First, children made more anticipatory switches overall in English videos, compared to non-English videos. This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior, consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}. Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch if they can expect an immediate response from the addressee. If so, then children's attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).

@@

Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}). This effect is straightforward: longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.

Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}). Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected. Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p} $=$ .053; Figure \ref{fig:E1-AduTs}). Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}). This interaction suggests that adults were less likely overall to make anticipatory switches at non-questions (as is evident for adults and children in Figure \ref{fig:E1-randvsreal}), and so did not benefit from extra time to do so compared to long gaps for questions.

@@

Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions. Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}). Prior work using online, metalinguistic tasks has shows that participants can use linguistic cues to accurately predict upcoming turn ends. The current results suggest that, in their spontaneous predictions about third-party conversation, both children and adults monitor the linguistic structure of unfolding turns for cues to upcoming responses.

>>>>

\usepackage{graphicx}
\usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{subfig}
\usepackage{tipa}
\usepackage{float}
\usepackage{longtable}
\usepackage[section]{placeins}

@@

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking: Early non-verbal interactions likely establish the protocol by which children come to use language with others. How do children integrate linguistic knowledge with these preverbal turn-taking abilities?

@@

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. Conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!'').

@@

@@

The majority of work on children's early turn taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions: They take every look, vocalization, arm flail, and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern. Infants' contingent vocalizations in the first few months of life show very fast timing (though with a lot of vocal overlap) that, by nine months, slows down considerably, only gradually speeding up again after 12 months \citep{hilbrink2015}. Taking turns with brief transitions between speakers is difficult for children; while their avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next few years \citep{casillas2016, garvey1984, ervin-tripp1979}. This puzzling pattern is likely due to their linguistic development: Taking turns on time is easier when the response is a simple vocalization rather than a linguistic utterance. Integrating language into the turn-taking system may be one major factor in children's delayed responses \citep{casillas2016}.

While children, like adults, could use linguistic cues in the ongoing turn to make predictions about upcoming turn structure, studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions. Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs. trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}. Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months they can start using prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}. In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.

%If response planning (i.e., language production) is the primary hurdle in young children's spontaneous turn taking, we should find evidence that children understand turn-taking behaviors before they are able to produce the behaviors themselves. This hypothesis has been recently explored in experimental settings, but results are mixed. One study found that 12-month-olds make more predictive gaze shifts to a responder while watching human verbal conversation compared to conversation-like interactions with objects \citep{bakker2011}, but another only found a similar effect at 36 months \citep{hofsten2009}. However, neither of these two studies had baselines to which the turn-relevant looking behavior could be compared. A baseline measurement is critical because there may be developmental differences in gaze shifting between conversational participants, even if the shifting is not related to turn structure. Such developmental differences could produce artifactual changes in measures of turn-contingent shifting.

@@

Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure. They asked 6-, 12-, 24-, 36-month-old, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change. They showed their participants two types of conversation videos---one normal and one with flattened pitch (i.e., with flattened intonation contours)---to test the role of intonation in participants' anticipatory predictions about upcoming speech. Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance. Among those, only 36-month-olds were affected by a lack of intonation contours, leading Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically. They also suggest that intonation takes a secondary role in turn prediction, following their and other studies findings with adults (e.g., \citealp{de-ruiter2006}), but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 36 months in their data).

Although the Keitel et al. \citeyearpar{keitel2013} study constitutes a substantial advance over previous work in this domain, it has its own limitations. Because these limitations directly inform our own study design, we review them in some detail. First, their estimates of baseline gaze frequency (``random'' in their terminology) were not random. Instead, they used gaze switches during ongoing speech as a baseline. Ongoing speech is perhaps the period in which switching is least likely to occur \citep{hirvenkari2013}, thus maximizing chances of finding a difference between gaze frequency at turn transitions and their baseline rate. A more conservative baseline would be to compare participants' looking behavior at turn transitions to their looking behavior during randomly selected windows of time throughout the stimulus, including turn transitions. We follow this conservative approach in our work.

@@

Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development. We present two experiments in which we measure children's anticipatory gaze to responders while watching conversation videos with natural (people using English vs. non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues. We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment.

Our findings suggest that children and adults use linguistic cues to make predictions about upcoming turn structure, but that they primarily do so to predict speaker transitions after a question. We also find that children make more predictions than  expected by chance, even at age XX, but that it takes several years before they recruit linguistic cues to anticipate responses more often after questions. We found no direct evidence of an early prosody advantage in children's anticipations and, further, no evidence that prosodic or lexical cues alone can substitute for their combination in the full linguistic signal, as is proposed for adults \citep{de-ruiter2006}.

@@

After recording, we combined the audio and video files by hand, and cropped each recording to the 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language condition (English vs. non-English), inter-turn gap duration (in milliseconds), and speech act (question vs. non-question). The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't: We asked five native American English speakers to listen to the audio signal for each turn and judge whether it sounded like a question. We then coded turns with at least 80\% ``yes'' responses as questions.

@@

Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively). They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: They gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}). Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not. Children looked at the current speaker less than adults did during the non-English videos. Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure. Overall, participants' pattern of gaze to current speakers demonstrated that they performed basic turn tracking during the videos, regardless of language. Figure \ref{fig:E1-randvsreal} shows participants' anticipatory gaze rates across age, language condition, and transition type, in the real data and the random baseline (described below).

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/E1-samples-by-lang-groups-trans-types.png}
\end{center}
\caption{Anticipatory gaze rates across language condition and transition type for the real (red and blue) and randomly permuted baseline (gray). Vertical bars represent the standard error.} 
\label{fig:E1-randvsreal}
\end{figure}

@@

We identified anticipatory gaze switches for all 36 usable turn transitions, based on the criteria outlined in Section \ref{sec:algorithm}, and analyzed them for effects of language, transition type, and age with two mixed-effects logistic regressions \citep{lme4, R}. We built one model each for children and adults. We modeled children and adults separately because effects of age are only pertinent to the children's data. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors. Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}. We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age (3, 4, 5; numeric), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report are all qualitatively unchanged by the exclusion of their random slopes. We have left the random slopes in because of minor participant-level variation in the predictors modeled.} The adult model included condition, transition type, duration, and their interactions as predictors with participant and item included as random effects and random slopes of condition, transition type, and their interaction for participant.

@@

We estimated the probability that these patterns were the result of random looking by running the same regression models on participants' real eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows. This process involved: (1) Randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is defined in Figure \ref{fig:criterion}), thereby randomly redistributing the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described in Section \ref{sec:algorithm}), this time using the randomly permuted analysis windows, and (3) modeling the anticipatory gazes from the randomly permuted data with the same statistical models we used for the original data (Section \ref{sec:models1}; Table \ref{tab:E1-models}). Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each random permutation.

\begin{figure}[ht!]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of analysis window permutations for a stimulus with five turn transitions. The windows were $\pm$300 msec around the inter-turn gap.} 
\label{fig:shuffling}
\end{figure}

@@

We completed this baseline procedure on 5,000 random permutations of the original turn transition analysis windows and compared the \textit{t}-values from each predictor in the original models (Table \ref{tab:E1-models}) to the distribution of \textit{t}-values for each predictor in the 5,000 models of the randomly permuted datasets.\footnote{We report \textit{t}-values rather than beta estimates because the standard errors in the randomly permuted data models were much higher than for the original data. For those interested, plots of the beta and standard error distributions are available in the Supplementary Materials.} We could then test whether significant effects from the original statistical models differed from the random baseline by calculating the proportion of random data \textit{t}-values exceeded by the original \textit{t}-value for each predictor, using the absolute value of all \textit{t}-values for a two-tailed test. For example, children's original ``language condition'' \textit{t}-value was $|$3.429$|$, which is greater than 99.9\% of all $|$\textit{t}-value$|$ estimates from the randomly-permuted data models (i.e., \textit{p}$=$.001). This leads us to conclude that that the effect of language condition in the original model was highly unlikely to be the result of random gaze shifting.

We excluded the output of random-permutation models that gave convergence warnings in order to remove unreliable estimates from our analyses (non-converging models were 22.4\% and 24.4\% of all models for children and adults, respectively; see the Supplementary Materials for more information on model exclusion).

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking (see Figures \ref{fig:E1-ChiTs} and \ref{fig:E1-AduTs} for the full distributional data). For the children's data, the original \textit{t}-values for language condition, gap duration, the age-language condition interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{t}-values for the randomly permuted data (99.9\%, 95.5\%, 99.4\%, and 96\%, respectively). Similarly, the adults' data showed significant differentiation from the randomly permuted data for two of the three originally significant predictors---transition type and the transition type-gap duration interaction (greater than 99.9\% and 99.7\% of random \textit{t}-values, respectively)---with marginal differentiation for the interaction of language condition and transition type (greater than 94.7\% of random \textit{t}-values).

\paragraph{Developmental effects}
The original statistical model of the children's real data revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was highly unlikely to have come from random looking (Figure \ref{fig:E1-randvsreal}). To further explore this effect, we compared the average effect of language condition for each age group: We extracted the average difference score for the two language conditions (English minus non-English) for each subject, computing an overall average for each random permutation of the data. For each random permutation, we then made pairwise comparisons of the average difference scores across pariticipant age groups, finding that, while 3- and 4-year olds showed similarly large effects of language condition, 5-year-olds showed a significantly smaller effect of language condition, compared to both younger age groups. In other words, the difference in the effect of language condition for 5-year-olds compared to younger children was larger than would be expected by chance in 99.52\% of the randomly permuted data sets for 3-year-olds and 99.96\% of the data sets for 4-year-olds---differences of \textit{p}$<$.01 and \textit{p}$<$.001, respectively (see Figure \ref{fig:E1-lgageinteraction} for difference score distributions).

When does spontaneous turn prediction emerge developmentally during natural speech? To test whether the youngest age group (3-year-olds) already exceeded chance in their anticipatory gaze switches, we used two-tailed \textit{t}-tests to compare their real gaze rates to the random baseline in the most familiar speech condition for our participants: English. We found that three-year-olds made anticipatory gaze switches significantly above chance, both when all transitions were considered (\textit{t}(22.824)$=$-4.147, \textit{p}$<$.001) and for question transitions alone (\textit{t}(21.677)$=$-5.268, \textit{p}$<$.001).

@@

Language condition (English vs. non-English) affected children's anticipations in two ways (Table \ref{tab:E1-models};  Figure \ref{fig:E1-randvsreal}). First, children made more anticipatory switches overall in English videos, compared to non-English videos. This effect suggests that lexical access is important for children's ability to anticipate upcoming turn structure; children had no lexical access to the speech in the non-English videos, though they did have access to (non-native) prosodic cues and non-verbal behavior. This finding is consistent with prior work on turn-end prediction in adults (\citealp{de-ruiter2006}; \citealp{magyari2012}) and children \citep{keitel2013}. Second, children systematically made more anticipatory switches after hearing a question compared to a non-question, but only in the English condition, suggesting that, when children have access to lexical cues, they are more likely to make an anticipatory gaze switch if they can expect an immediate response from the addressee. If so, then children's (and adults') attention to lexical cues for turn taking may primarily be in monitoring the signal for cues to questionhood (e.g., subject-auxiliary inversion, \textit{wh}-words, etc.).

@@

Finally, children showed an effect of gap duration (Table \ref{tab:E1-models}). This effect is straightforward: Longer gaps resulted in longer analysis windows, yielding more time for children to make an anticipatory gaze.

Adults' anticipatory gaze rates were also affected by transition type, language condition, and gap duration (Table \ref{tab:E1-models}), none of which could be easily explained by a baseline of random gaze switching (Figure \ref{fig:E1-AduTs}). Like children, adults made more anticipatory switches after hearing questions compared to non-questions, suggesting that anticipation mattered more to them when an immediate response was expected. Also like children, the advantage for questions was driven by lexical access such that adults must have relied on lexicosyntactic cues to questionhood in picking out turns that potentially require an immediate response, though this effect was only marginally divergent from the distribution of randomly permuted data (\textit{p}$=$.053; Figure \ref{fig:E1-AduTs}). Finally, adults' anticipation rates were also affected by gap duration, but more so for questions than non-questions (Table \ref{tab:E1-models}), suggesting that adults were less likely overall to make switches at non-questions, and so did not benefit from extra time to do so.

@@

Interestingly, adults and children both were strongly affected by transition type, in that they made more anticipatory switches after hearing questions, compared to non-questions. Even in the English videos, when participants had full access to linguistic cues, their rates of anticipation were relatively low---in fact, comparable to the non-English videos---unless the turn was a question (Figure \ref{fig:E1-randvsreal}). Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari12, de-ruiter2006}. The current results add a new dimension to our understanding of how listeners make predictions about turn ends: Both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to upcoming responses.
